Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-11-19 ms/spaceness.tex
correcting supp fig references

An important step in creating any spatial model is dealing with range edges. 
Because local population density is used to model competition, edge or corner populations can be assigned artificially high fitness values 
because they lack neighbors within their interaction radius but outside the bounds of the simulation.
We approximate a decline in habitat suitability near edges by decreasing the probability of survival proportional to the square root of distance to edges in units of $\sigma$. 
The final probability of survival for individual $i$ is then
\begin{equation}
    s_{i} = p_{i} 
        \min(1,\sqrt{x_{i}/\sigma})
        \min(1,\sqrt{y_{i}/\sigma})\\
        \min(1,\sqrt{(W - x_{i})/\sigma})
        \min(1,\sqrt{(W - y_{i})/\sigma)}
\end{equation}
where $x_i$ and $y_i$ are the spatial coordinates of individual $i$, and $W$ is the width (and height) of the square habitat.
This buffer roughly counteracts the increase in fitness individuals close to the edge would otherwise have, though the effect is relatively subtle \ref{fig:edge_scaling}.

@@

Figure \ref{fig:sumstats} shows the effect of varying neighborhood size and sampling strategy on the site frequency spectrum (Figure~\ref{fig:sumstats}A) and several standard population genetic summary statistics (Figure~\ref{fig:sumstats}B). Consistent with findings in island and stepping stone simulations \citep{Stadler2009}, the SFS shows a significant enrichment of intermediate frequency variants in comparison to the nonspatial expectation. This bias is most pronounced below neighborhood sizes $\leq 100$ and is exacerbated by midpoint and point sampling of individuals (depicted in Figure~\ref{fig:samplemap}). 
Reflecting this, Tajima's $D$ is quite positive in the same situations (Figure~\ref{fig:sumstats}B). Notably, the point at which Tajima's $D$ approaches 0 differs strongly across sampling strategies -- varying from a neighborhood size of roughly 50 for random sampling to at least 1000 for midpoint sampling.

@@

\begin{figure*}[p]
\centering
\includegraphics[]{figures/slim_v_msp_sfs.pdf}
\caption{Site frequency spectra from a simulation with neighborhood size = 12.5 when mutations are recorded directly in SLiM (blue line) or applied later in msprime (red line).}
\label{fig:slim_v_msp_sfs2} 
\end{figure*}
\FloatBarrier

>>>>

An important step in creating any spatial model is dealing with range edges. 
Because local population density is used to model competition, edge or corner populations can be assigned artificially high fitness values 
because they lack neighbors within their interaction radius but outside the bounds of the simulation.
We approximate a decline in habitat suitability near edges by decreasing the probability of survival proportional to the square root of distance to edges in units of $\sigma$. 
The final probability of survival for individual $i$ is then
\begin{equation}
    s_{i} = p_{i} 
        \min(1,\sqrt{x_{i}/\sigma})
        \min(1,\sqrt{y_{i}/\sigma})\\
        \min(1,\sqrt{(W - x_{i})/\sigma})
        \min(1,\sqrt{(W - y_{i})/\sigma)}
\end{equation}
where $x_i$ and $y_i$ are the spatial coordinates of individual $i$, and $W$ is the width (and height) of the square habitat.
This buffer roughly counteracts the increase in fitness individuals close to the edge would otherwise have, though the effect is relatively subtle (Figure \ref{fig:edge_scaling}).

@@

Figure \ref{fig:sumstats} shows the effect of varying neighborhood size and sampling strategy on the site frequency spectrum (Figure~\ref{fig:sumstats}, Figure \ref{fig:allsfs}) and several standard population genetic summary statistics (Figure~\ref{fig:sumstats}B; additional statistics are shown in Figure \ref{fig:allsumstats}). Consistent with findings in island and stepping stone simulations \citep{Stadler2009}, the SFS shows a significant enrichment of intermediate frequency variants in comparison to the nonspatial expectation. This bias is most pronounced below neighborhood sizes $\leq 100$ and is exacerbated by midpoint and point sampling of individuals (depicted in Figure~\ref{fig:samplemap}). 
Reflecting this, Tajima's $D$ is quite positive in the same situations (Figure~\ref{fig:sumstats}B). Notably, the point at which Tajima's $D$ approaches 0 differs strongly across sampling strategies -- varying from a neighborhood size of roughly 50 for random sampling to at least 1000 for midpoint sampling.

@@

\begin{figure*}[p]
\centering
\includegraphics[]{figures/slim_v_msp_sfs.pdf}
\caption{Site frequency spectra from a simulation with neighborhood size = 12.5 when mutations are recorded directly in SLiM (blue line) or applied later in msprime (red line).}
\label{fig:slim_v_msp_sfs} 
\end{figure*}
\FloatBarrier
